### 2019-07-04 article.tex
refresh reviewers' comments

% ------------------------------------------
%  IEEE ARTICLE TEMPLATE
% ------------------------------------------
% Author(s):
% Institution(s):
% ------------------------------------------
\documentclass[journal,twocolumn,letterpaper,10pt]{ieee-sty/IEEEtran}
\usepackage[acronym]{glossaries}

\usepackage{epsfig}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{blindtext}
\usepackage{xcolor}
\usepackage[ampersand]{easylist}
\usepackage[]{algorithm2e,algorithmicx}
\usepackage[bottom]{footmisc}
\usepackage{graphicx}

\usepackage{hyperref}

\renewcommand{\sectionautorefname}{\S}
\renewcommand{\subsectionautorefname}{\S}
\renewcommand{\subsubsectionautorefname}{\S}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

\usepackage{enumitem}
\usepackage{flushend}

% Generates the acronyms list
\makeglossaries

% Acronyms list
\input{acronyms/acronyms.tex}

% ------------------------------------------
%  MAIN DOCUMENT
% ------------------------------------------
\begin{document}

% Variables file
\input{variables.tex}

% Article title
\title{ Establishing Trusted I/O Paths for Client \\ Shielded Execution Systems with Aurora }

% Author(s) Name
\author{Hongliang~Liang,~\IEEEmembership{Member,~IEEE,}
	Mingyu Li,~Yixiu Chen,~Lin Jiang,~Zhuosi Xie,~Tianqi Yang\\
	%\{hliang, maxul, chenyixiu, jianglin, xiezhuosi, yangtianqi\}@bupt.edu.cn\\
	%Trusted Software and Intelligent System Lab\\
	%Beijing University of Posts and Telecommunications
	\thanks{H. Liang, M. Li, Y. Chen, L. Jiang, Z. Xie, T. Yang are with Trusted Software and Intelligent System Lab, Beijing University of Posts and Telecommunications, Beijing 100876, China. e-mail: (\{hliang, maxul, chenyixiu, jianglin, xiezhuosi, yangtianqi\}@bupt.edu.cn) }% <-this % stops a space
    \thanks{This work was supported by the National Natural Science Foundation of China (NSFC) under Grant No. U1713212 and 91418206.}
}

\maketitle

% Abstract
\input{abstract/abstract.tex}

% Keywords
\input{keywords/keywords.tex}

% Sections
\input{sections/main.tex}

% Bibliography
\bibliographystyle{IEEEtran}
% Bibliography file
\bibliography{bibliography/article}

\end{document}